---
title: "Betke-Briana-ADA-Homework-2"
author: "Briana Betke"
date: "3/9/2020"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 2
This is an Rmarkdown of my work for challenges 1-4 of homework 2. 

### Challenge 1: 
#### Step 1: Read in the necessary packages and data
```{r}
library(tidyverse)
library(cowplot)
theme_set(theme_cowplot()) # b/c no grids in my graphs!!
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)
```

#### Step 2. Filter out movies b/w 1920-1979 and runtime of < 240 minutes. Then create a variable called decates for each decade of movies.
```{r}
s <- filter(d, runtimeMinutes < 240 & startYear %in% 1920:1979) %>% 
  mutate(decade = if_else(startYear %in% 1920:1929,"20s",
                         if_else(startYear %in% 1930:1939,"30s",
                                 if_else(startYear %in% 1940:1949,"40s",
                                         if_else(startYear %in% 1950:1959,"50s",
                                                if_else(startYear %in% 1960:1969,"60s","70s"))))))

```

#### Step 3. Graph runtimes by decade
```{r}
# create histograms of runtimes by decade.  
p <- ggplot(s,aes(x=runtimeMinutes))+
  geom_histogram()+
  facet_wrap(~decade)
p
```

#### Step 4. Calculate mean and standard deviations by decade and save to results.
```{r}
results <- group_by(s, decade)%>%
  summarise(n_cases = n(),
            avg = mean(runtimeMinutes, na.rm = TRUE),
            popSD = sqrt(sum((runtimeMinutes - mean(runtimeMinutes))^2) / length(runtimeMinutes)))
results
```

#### Step 5. Single Sample
```{r}
Sample_decades <- group_by(s, decade) %>% sample_n(100) 

outcome <- group_by(Sample_decades, decade) %>%
  summarize(
    avg = mean(runtimeMinutes, na.rm = TRUE),
    SD = sd(runtimeMinutes, na.rm = TRUE),
    se = SD/sqrt(100) # not sure if this part it correct.....
  )
outcome
```

#### Step 6. 10000 samples of size 100





### Challenge 2:
#### Step 1. Calculate the probability that she will hear 13 or fewer calls during any given session.
```{r}
ppois(13, lambda = 18)
```

#### Step 2. Calculate the probability that she will hear no calls in a session.
```{r}
dpois(0, lambda = 18)
```

#### Step 3. Calculate the probability that she will hear exactly 7 calls in a session.
```{r}
dpois(7, lambda = 18)
```

#### Step 4. Calculate the probability that she will hear more than 20 calls in a session.
```{r}
ppois(20, lambda = 18, lower.tail = FALSE)
```

#### Step 5. Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 40.
```{r}
# plot probability mass function 
library(mosaic)
l <- 18
pmf_poisson <-
  plotDist(
    "pois",
    lambda = l,
    xlim = c(0,40),
    main = paste0("Poisson Distribution\nwith lambda=", l),
    xlab = "x",
    ylab = "Pr(X=x)"
  )
pmf_poisson
```

#### Step 6. Simulate 520 results from this distribution and plot the distribution using the historam() from mosaic.
```{r}
q <- rpois(520, lambda = 18)

histogram(q,  xlim = c(0,40))
```
The shape between the simulation and the probability mass funtion look similar. 

### Challenge 3:
#### Step 1. Read in the zombie data
```{r}
# get that zombie data 
f <- "https://raw.githubusercontent.com/difiore/ADA-datasets/master/zombies.csv"
d <- read_csv(f, col_names = TRUE)
```

#### Step 2. Calculating population mean and sd for each quatitative random variable
```{r}
pop_var <- function(x) {
  sum((x - mean(x))^2) / (length(x))
}

pop_sd <- function(x) {
  sqrt(pop_var(x))
}

results2 <- summarise(d,
  avgH = mean(height, na.rm = TRUE),
  avgW = mean(weight, na.rm = TRUE),
  avgZ = mean(zombies_killed, na.rm = TRUE),
  avgE = mean(years_of_education,na.rm = TRUE),
  avgA = mean(age,na.rm = TRUE),
  sdH = pop_sd(height),
  sdW = pop_sd(weight),
  sdZ = pop_sd(zombies_killed),
  sdE = pop_sd(years_of_education),
  sdA = pop_sd(age)
)
results2
```

#### Step 3. 
```{r}
# scatter plot of weight and height
# SET UP THE NEATO GRAPHS
p1 <- ggplot(data = d, aes(
  x = height,
  y = age,)) +
  geom_point(na.rm = TRUE)

p2 <- ggplot(data = d, aes(
  x = weight,
  y = age,)) +
  geom_point(na.rm = TRUE)

plot_grid(p1, p2, labels = c("A", "B"), label_size = 12, nrow = 1)
```

#### Step 4. 
